---
title: "final_project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Importing data and setting aside Test
```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(caret)
library(ROCR)

new_orleans_file <- "hp256wp2687_la_new_orleans_2019_08_13.csv.zip"

new_orleans <- 
#  read_csv(new_orleans_file)

set.seed(1)
categories <- sample(1:2, size = nrow(new_orleans), replace = TRUE, prob = c(0.8, 0.2))
train_uncleaned <- new_orleans[categories == 1,]
# test_uncleaned <- new_orleans[categories == 2,]

```


#Data cleaning and splitting into Train and Valid
```{r}
train_cleaned <- 
  train_uncleaned %>% 
  filter(year(date) >= 2012) %>% 
  mutate(
    time_of_day = 
      case_when(
        hour(time) < 6 ~               "night", 
        hour(time) >= 6 & hour(time) < 12 ~   "morning", 
        hour(time) >= 12 & hour(time) < 18 ~  "afternoon", 
        hour(time) >= 18 ~             "evening"
      )
  ) %>% 
  select(
    arrest_made, date, time_of_day, district, subject_age, subject_race, subject_sex, 
    officer_assignment, type, contraband_found, frisk_performed, search_person, 
    search_vehicle, search_basis, reason_for_stop
  ) %>% 
  group_by(reason_for_stop) %>% #cleaning out entries with more than one reason for stop
  filter(n() > 100) %>% 
  ungroup() %>% 
  group_by(officer_assignment) %>% #cleaning out entries with more than one officer assignment
  filter(n() > 5) %>% 
  ungroup() %>% 
  mutate(
    officer_assignment = 
      recode(
        officer_assignment, 
        "FOB" = "Other", 
        "ISB" = "Other",
        "MSB" = "Other", 
        "NCIC" = "Other", 
        "PIB" = "Other", 
        "Reserve" = "Other", 
        "SOD" = "Other", 
        "Superintendent" = "Other"
      ), 
    district = as.factor(district), 
    contraband_found = as.character(contraband_found), 
    contraband_found = if_else(is.na(contraband_found), "Not searched", contraband_found), 
    search_basis = as.character(search_basis),
    search_basis = if_else(is.na(search_basis), "Not searched", search_basis), 
    arrest_made = if_else(arrest_made == TRUE, 1, 0),
    month = month(date), 
    weekday = wday(date, label = TRUE), 
    weekday = if_else(weekday %in% c("Sun", "Sat"), "Weekend", "Weekday")
  ) %>% 
  filter_all(~ !is.na(.)) %>% 
  select(-date)

set.seed(1)
categories_2 <- sample(1:2, size = nrow(train_cleaned), replace = TRUE, prob = c(0.8, 0.2))
orleans_train <- train_cleaned[categories_2 == 1,]
orleans_valid <- train_cleaned[categories_2 == 2,]
```


#Data Exploration
```{r}
orleans_train %>% 
  count(arrest_made, subject_race) %>% 
  ggplot(aes(arrest_made, subject_race, size = n)) + 
  geom_point()
```


```{r}
chi_arrest <- function(dep_variable) {
  dep_variable = enquo(dep_variable)
  
  data_for_chi <- 
    orleans_train %>% 
    count(arrest_made, !!dep_variable) %>% 
    spread(key = arrest_made, value = n) %>% select(-!!dep_variable)
  
  chi <- chisq.test(data_for_chi, correct = FALSE)
  
  v <- sqrt(chi$statistic / sum(data_for_chi))
  
  return(c(round(chi$p.value, 2), round(v, 2), str_c(dep_variable)))
}

as.tibble(
  chi_arrest(time_of_day) %>% 
  rbind(chi_arrest(district)) %>% 
  rbind(chi_arrest(subject_race)) %>% 
  rbind(chi_arrest(subject_sex)) %>% 
  rbind(chi_arrest(officer_assignment)) %>% 
  rbind(chi_arrest(type)) %>% 
  rbind(chi_arrest(contraband_found)) %>% 
  rbind(chi_arrest(frisk_performed)) %>% 
  rbind(chi_arrest(search_person)) %>% 
  rbind(chi_arrest(search_vehicle)) %>% 
  rbind(chi_arrest(search_basis)) %>% 
  rbind(chi_arrest(reason_for_stop))
) %>% 
  select(
    variable = V4, 
    p_value = V1, 
    v = `X-squared`
  )

```


```{r}
orleans_train %>% 
  filter(!is.na(contraband_found)) %>% 
  ggplot(aes(contraband_found, arrest_made)) + 
  geom_count() + 
  scale_size(range = c(1, 20), breaks = seq(0, 30000, 10000)) + 
  labs(
    x = "Search Basis", 
    y = "Arrest Made", 
    title = "Distribution of Arrest Made versus Contraband Found"
  )

```



###Boxplots for Age
```{r}
age_boxplots <- function(variable) {
  variable = enquo(variable)
  
  orleans_train %>% 
    filter(!is.na(!!variable)) %>% 
    ggplot(aes(!!variable, subject_age)) + 
    geom_boxplot()
}

# age_boxplots(arrest_made)
# age_boxplots(time_of_day)
# age_boxplots(district)
# age_boxplots(subject_race)
# age_boxplots(subject_sex)
# age_boxplots(officer_assignment)
# age_boxplots(type)
# age_boxplots(contraband_found)
# age_boxplots(frisk_performed)
# age_boxplots(search_person)
# age_boxplots(search_vehicle)
# age_boxplots(search_basis)
# age_boxplots(reason_for_stop)

orleans_train %>% 
  filter(!is.na(subject_race)) %>% 
  ggplot(aes(subject_race, subject_age)) + 
  geom_boxplot() + 
  labs(
    x = "Subject Race", 
    y = "Subject Age", 
    title = "Distribution of Age by Race"
  )
```

#Model Training for arrest_made

###Train and Valid with Logistic Regression
```{r}
logit_model <- glm(arrest_made ~ ., family = binomial(link = 'logit'), data = orleans_train)

orleans_valid <- 
  orleans_valid %>% 
  mutate(
    pred = predict(logit_model, .), 
    pred_clean = as.factor(if_else(pred > 0, 1, 0)),
    arrest_made = as.factor(arrest_made)
  )
  
  
Y <- orleans_valid %>% pull(arrest_made)
pred <- orleans_valid %>% pull(pred_clean)

#Precision
posPredValue(pred, Y, positive = "1")

#recall
sensitivity(pred, Y, positive = "1")


confusionMatrix(pred, Y)
```


###Cross Validation for Logistic Regression
```{r}
ctrl <- trainControl(method = "repeatedcv", number = 10, savePredictions = TRUE)

train.logit <- 
    train(
      arrest_made ~ ., 
      data = train_cleaned %>% mutate(arrest_made = as.factor(arrest_made)),
      method = "glm",
      family = "binomial",
      trControl = ctrl, 
      tuneLength = 5
    )

# Predict
pred_cv <- predict(train.logit, train_cleaned)

# Produce confusion matrix from prediction and data used for training
confusionMatrix(pred_cv, train.logit$trainingData$.outcome, mode = "everything")
```

###AUC and ROC for CV logistic
```{r}
labels <- train_cleaned %>% pull(arrest_made)

pred_cv_auc <- prediction(as.integer(pred_cv), as.integer(labels))

perf <- performance(pred_cv_auc, "auc")
auc <- perf@y.values[[1]]
auc

perf_roc <- performance(pred_cv_auc, "tpr", "fpr")
plot(perf_roc)
```
##Classification contd.
###Cross Validation using SVM
```{r}
ctrl <- trainControl(method = "repeatedcv", number = 10, savePredictions = TRUE)

train.svm <- 
    train(
      arrest_made ~ ., 
      data = train_cleaned %>% mutate(arrest_made = as.factor(arrest_made)),
      method = "svmLinear",
      trControl = ctrl,
    )
# Predict
pred_cv_svm <- predict(train.svm, train_cleaned)

# Produce confusion matrix from prediction and data used for training
confusionMatrix(pred_cv_svm, train.svm$trainingData$.outcome, mode = "everything")
```

###Cross Validation using KNN
```{r}
ctrl <- trainControl(method = "cv", number = 10, savePredictions = TRUE)

train_cleaned_temp <- dummyVars(" ~ .", data = train_cleaned %>% select(-arrest_made), fullRank=T) 
train_cleaned_dummified <- data.frame(predict(train_cleaned_temp, newdata = train_cleaned  %>% select(-arrest_made)))
train_cleaned_dummified$arrest_made = train_cleaned$arrest_made

train.knn <- 
    train(
      arrest_made ~ ., 
      data = train_cleaned_dummified %>% mutate(arrest_made = as.factor(arrest_made)),
      method = "knn",
      trControl = ctrl,
      tuneLength = 5,
      
    )
# Predict
pred_cv_knn <- predict(train.knn, train_cleaned_dummified)


# Produce confusion matrix from prediction and data used for training
confusionMatrix(pred_cv_knn, train.knn$trainingData$.outcome, mode = "everything")
```


###Cross Validation using RF
```{r}
#metric <- "ROC"  
ctrl <- trainControl(method = "cv", number = 10, savePredictions = TRUE)
#rf_default <- train(Class~., data=dataset, method="rf", metric=metric, tuneGrid=tunegrid, trControl=control)


train.rf <- 
    train(
      arrest_made ~ ., 
      data = train_cleaned_dummified %>% mutate(arrest_made = as.factor(arrest_made)),
      method = "rf",
      trControl = ctrl,
      metric="ROC",
      tuneLength = 5,
    )
# Predict
pred_cv_rf <- predict(train.rf, train_cleaned)


# Produce confusion matrix from prediction and data used for training
confusionMatrix(pred_cv_ef, train.rf$trainingData$.outcome, mode = "everything")
```

###Continuous variable prediction 
```{r}
train.lasso <- 
    train(
      arrest_made ~ ., 
      data = train_cleaned %>% mutate(arrest_made = as.factor(arrest_made)),
      method = "glm",
      family = "binomial",
      trControl = ctrl, 
      tuneLength = 5
    )

# Predict
pred_cv <- predict(train.logit, train_cleaned)

# Produce confusion matrix from prediction and data used for training
confusionMatrix(pred_cv, train.logit$trainingData$.outcome, mode = "everything")
```




#Predicting Continuous Variable

```{r}
model_ols <- lm(subject_age ~ ., data = train_cleaned)

ctrl <- trainControl(method = "repeatedcv", number = 10, savePredictions = TRUE)

train.ols <- 
    train(
      subject_age ~ ., 
      data = train_cleaned,
      method = "lm",
      trControl = ctrl, 
      tuneLength = 5
    )

# Predict
pred_cv_ols <- predict(train.ols, train_cleaned)

train_cleaned %>% 
  cbind(pred_cv_ols) %>% 
  mutate(error = (subject_age - pred_cv_ols)^2) %>% 
  summarize(rmse = sqrt(mean(error, na.rm = TRUE)))
```

### Regression Cross Validation using KNN
```{r}
ctrl <- trainControl(method = "cv", number = 10, savePredictions = TRUE)

train_cleaned_temp_reg <- dummyVars(" ~ .", data = train_cleaned %>% select(-subject_age), fullRank=T) 
train_cleaned_dummified_reg <- data.frame(predict(train_cleaned_temp_reg, newdata = train_cleaned  %>% select(-subject_age)))
train_cleaned_dummified_reg$subject_age = train_cleaned$subject_age

train.knn <- 
    train(
      subject_age ~ ., 
      data = train_cleaned_dummified_reg,
      method = "knn",
      trControl = ctrl,
      metric = "RMSE",
      tuneLength = 5,
      
    )
# Predict
pred_cv_knn <- predict(train.knn, train_cleaned_dummified)


# Produce confusion matrix from prediction and data used for training
confusionMatrix(pred_cv_knn, train.knn$trainingData$.outcome, mode = "everything")
```

###Regression with CART 
```{r}
ctrl <- trainControl(method = "cv", number = 10, savePredictions = TRUE)

train_cleaned_temp_reg <- dummyVars(" ~ .", data = train_cleaned %>% select(-subject_age), fullRank=T) 
train_cleaned_dummified_reg <- data.frame(predict(train_cleaned_temp_reg, newdata = train_cleaned  %>% select(-subject_age)))
train_cleaned_dummified_reg$subject_age = train_cleaned$subject_age

train.rpart_reg <- 
    train(
      subject_age ~ ., 
      data = train_cleaned_dummified_reg,
      method = "rpart",
      trControl = ctrl,
      metric = "RMSE",
      tuneLength = 5,
      
    )
# Predict
pred_cv_rpart_reg <- predict(train.rpart_reg, train_cleaned_dummified)


# Produce confusion matrix from prediction and data used for training
confusionMatrix(pred_cv_rpart_reg, train.rpart_reg$trainingData$.outcome, mode = "everything")
```


#Regression with RF
```{r}
ctrl <- trainControl(method = "cv", number = 10, savePredictions = TRUE)

train_cleaned_temp_reg <- dummyVars(" ~ .", data = train_cleaned %>% select(-subject_age), fullRank=T) 
train_cleaned_dummified_reg <- data.frame(predict(train_cleaned_temp_reg, newdata = train_cleaned  %>% select(-subject_age)))
train_cleaned_dummified_reg$subject_age = train_cleaned$subject_age

train.rf_reg <- 
    train(
      subject_age ~ ., 
      data = train_cleaned_dummified_reg,
      method = "rf",
      trControl = ctrl,
      metric = "RMSE",
      tuneLength = 5,
      
    )
# Predict
pred_cv_rf_reg <- predict(train.rf_reg, train_cleaned_dummified)


# Produce confusion matrix from prediction and data used for training
confusionMatrix(pred_cv_rf_reg, train.rf_reg$trainingData$.outcome, mode = "everything")
```





response variable
- arrest_made

Covariates
- date
- time 
- district
- subject_age
- subject_race
- subject_sex
- officer_assignment
- type
- contraband_found (would be problematic if the outcome is arrest)
- frisk_performed 
- search_person
- search_vehicle
- search_basis
- reason_for_stop (but clean the data)

Keep for later
- vehicle_color, vehicle_model, vehicle_year 
- outcome (Turn NA into No action taken)







